Gait Monitoring using Wireless Sensors\\ \vspace{4mm} \large{Final Report 18-748}\\ \large{Spring 2017}

### 2017-05-10 paper.tex
Added results section and a bundle of figures

\section{Results}
\

>>>>

\section{Results}
Our results are best demonstrated through a series of the screenshots acquired during
multiple demos. They capture the objective of our project and demonstrate the value of the
current prototype as well as the points where improvements could be made. 
\subsection{Visualization}
Our final data visualization included a simple avatar and representations of the
orientation of all six slave nodes. Figure~\ref{fig:vis} shows the final visualization layout.
Both elements of the visualization moved in real time as data passed in from the master.
For the orientations, each node's axes move independent of the other nodes, so the
movements provide the user with a sense of the raw noise included in the orientation
measurements. If any of the node's IMUs reports some kind of spurious, outlying data
point, the user will observe a sudden change in that IMU's axis representation. Future
work would filter out sudden changes to prevent this phenomena.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.95\columnwidth]{figs/vis}
  \caption{{\bf Final visualization layout}}
  \label{fig:vis }
\end{figure}

\subsection{Gait parameter graphs}
In adition to visualizing a subject's movements, our system can generate graphs of a
subject's knee extension and stride length in real time. Figure~\ref{fig:extend} shows the
graph of knee extension on the right as our subject bends his knee. The graph contains
separate extension angles for the left and right knee, so it is simple to notice if the
range of motion in one knee is more restricted than the other.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.95\columnwidth]{figs/extend}
  \caption{{\bf Real time knee extension reporting}}
  \label{fig:extend }
\end{figure}

The stride length graph appears in nearly real time, but requires additional processing
of the segments of multiple data points to find where steps occur, thus delaying it
slightly behind real time. However, it still provides useful, prompt feedback on the
length of a subject's stride as shown in Figure~\ref{fig:length}. The graph shows the
distance between the right and left foot, so the rise and fall pattern observed in the
graph in this figure indicates the feet separating as a subject's leg moves forward and coming
back together as the subject swings one foot past the other on the next step.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.95\columnwidth]{figs/length}
  \caption{{\bf Stride length reporting}}
  \label{fig:length}
\end{figure}

\subsection{Fall Detection}
In testing our gait monitoring system, we found that the orientation reported by the nodes
strapped to a patient's thighs provided a reliable fall indicator for our subject. When
the subject is walking, the measurement is stable and indicates that the node is
perpendicular to the ground plane. If a subject falls, the orientation data indicates a
rapid change in the angle between the node's central axis and the ground as demonstrated
in Figure~\ref{fig:falling}. Though our system relies primarily on the obvious change in
the grph that a system operator can visually observe, an automatic mechanism based on a
set threshold for the angle between the nodes and the ground would suffice to warn
caregivers that a patient has fallen.

\begin{figure}[h]
  \centering
  \includegraphics[width=.99\columnwidth]{figs/falling}
  \caption{{\bf Angle reporting for fall detection }}
  \label{fig:falling}
\end{figure}
